---
title: "SCATS-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SCATS-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SCATS)
```

The primary goal of the following function is to determine the number of vehicles passing through each approach during the three designated peak hours—namely, "6 to 7 am" (AM), "1 to 2 pm" (MD), and "5 to 6 pm" (PM).

```{r load-packages, include=FALSE, cache=FALSE}
# Load the necessary library if not already loaded
library(dplyr)
library(readxl)
library(tidyr)
```

```{r message=FALSE, warning=FALSE, cache=FALSE}
# Select the relevant columns for summing
X6802 <- read_excel("6802.xlsx")
cols_to_sum <- c(10,17,21)

# Group by Approach and calculate the sum for the selected columns
results <- X6802 %>%
  group_by(Approach) %>%
  summarise(across(all_of(cols_to_sum), sum, na.rm = TRUE))

# Display the result
print(results)
```

Summarizing the data will yield the highest car count within a 15-minute interval, along with the number of NA values, facilitating an analysis of both the peak rate and instances of shortages over time.

```{r warning=FALSE}
summary(X6802)
```

The objective of the following code is to reorganize the data, reducing the number of columns from 28 to 8, thereby facilitating a more straightforward description and analysis.

```{r}
X6802N <- X6802 |>
#It selects all columns excluding those between "Junctions" and "Interval," transforming the data into a longer format. 
#In this new format, the variable names are moved into a new column named "hours" while the corresponding values are placed in a separate column labeled "vehicles"
  pivot_longer(cols = -c(Junctions:Interval),
               names_to = "hours",
               values_to = "vehicles")
```

