---
title: "SCATS-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SCATS-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SCATS)
data("X6802N")
```

```{r load-packages, include=FALSE, cache=FALSE}
# Load the necessary library if not already loaded
library(tidyverse)
```


The primary goal of the following function is to create a daily profile of the total number of vehicles that go through junction 6802. which will help in understanding the flow of vehicles and to determine the peak hours specific to this junction and if it is the same as the general network peak hours or not. Peak hours intervals are the following: "6 to 9 am" (AM Peak), "12 to 3 pm" (MD Peak), "5 to 8 pm" (PM Peak). Network peak hours are:  "6 to 7 am" (AM), "1 to 2 pm" (MD), and "5 to 6 pm" (PM).

```{r message=FALSE, warning=FALSE, cache=FALSE}

# Assigning X6802N data set into junction_data to work on
junction_data <- X6802N %>%
# AS the aim is to have a daily profile the data will be grouped passed on the hours form 0 to 23
  group_by(hours) %>%
# We sum vehicles into total_vehicles variable based on hours
  summarise(total_vehicles = sum(vehicles, na.rm = TRUE))
# As hours where not in order from 1 to 23. Here we convert "hours" to a factor to present it in the correct order
junction_data$hours <- factor(junction_data$hours, levels = 0:23)

# Here we plot the daily profile
ggplot(junction_data, aes(x = hours, y = total_vehicles)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Daily Profile (Sum of All Approaches)",
      x = "Hours",
      y = "Total Number of Vehicles")

```

Summarizing the data will yield the highest car count within a 15-minute interval, along with the number of NA values, facilitating an analysis of both the peak rate and instances of shortages over time.

```{r warning=FALSE}
summary(X6802N)
```


